Differences in Homicide Case Characteristics Indicates Why Justice is Not Served

### 2024-12-02 paper/paper.qmd
Finished data section and started working on idealized survey

The raw dataset we retrieved from The Washington Post using a script that downloads the CSV file from their GitHub repository contains 52179 observations with one observation being a homicide case. However, since we narrowed our scope to focus on the two of the most populated US cities, Los Angeles and Chicago, the number of observations in our cleaned dataset ended up being 2820. Our data look as follows:

@@

@tbl-dataset-preview and @tbl-dataset-statistics indicates our variables of interest, which are the following: victim_race, victim_age, victim_sex, city, disposition, year, month, and arrest_was_not_made. victim_race represents the race of the homicide victim, which can be "White" (248 observations), "Hispanic" (1242 observations), "Black" (1216 observations), "Asian" (58 observations), and "Other" (56 observations). victim_age signifies the age of the homicide victim at the time of their death and is defined as a whole number. @tbl-dataset-statistics reveals that the mean victim_sex is the sex of the homicide victim where they are either identified as a "female" or "male". The city variable defines the city the victim is reported to have been found in. The disposition variable is the specific status of a homicide case where a case can fall in either of the following three status: "Closed by arrest", "Open/No arrest", and "Closed without arrest". The year variable represents a year from 2010 to 2017 that indicates the year the homicide took place. Following this, the month variable represents the month a homicide took place. arrest_was_not_made is a variable that was constructed based on the disposition variable with "Closed by arrest" being converted to a 0 and "Open/No arrest" and "Closed without arrest" being converted to a 1. arrest_was_not_made indicates the status of a homicide case as either being unsolved, which is denoted by a 1, and solved, which is denoted by a 0.

@@

- TODO: A thorough discussion of measurement, relating to the dataset, is provided in the data section. Please ensure that you explain how we went from some phenomena in the world that happened to an entry in the dataset that you are interested in.

Some paragraphs about how we go from a phenomena in the world to an entry in the dataset.

- How homicide cases are recorded by the Chicago Police and Los Angeles Police 
- How The Washington Post obtain the data and what did they do

The Federal Bureau of Investigation (FBI) has a program called the Uniform Crime Reporting (UCR) Program to generate statistics for the public [@citeFBIinfo]. The FBI originally had a system under the UCR called Summary Reporting System (SRS), which obtained details about different crimes taking place from law enforcement agencies nationwide such as victim information [@citeFBIinfo]. However, the SRS was replaced with a new system called the National Incident-Based Reporting System (NIBR) in 2021 that obtained more details about various crimes [@citeFBIinfo]. For The Los Angeles Police Department, after a homicide case occurs, the investigating team handwrites information into physical crime reports with details like the type of crime, the premise the crime occured at, and the age and ethnicity of the victim [@citeLApolice]. The crime reports are then transcribed into a digital format and then sent to the SRS monthly [@citeLApolice].

## Outcome variable

## Predictor variables
The variables, disposition and month are not predictor variables in our model but they are used to investigate trends between homicide case information and the status of the homicide being solved or unsolved.

@@

## Model set-up
With our model, we will make the assumption that there is a relationship between homicide case information like the victim's race, victim's age, victim's sex, the city, and the year with a homicide case being unsolved. We also assume that the predictor variables are independent from one another, which we check in @sec-model-details using variance inflation factor (VIF) and it indicates the predictors are not highly correlated with each other. We define our model as follows:

\begin{align*}
y_i|\pi_i &\sim \mbox{Bern}(\pi_i) \\
\mbox{logit}(\pi_i) &= \beta_0 + \beta_1 \times \mbox{victim\_race}_i + \beta_2 \times \mbox{victim\_age}_i + \beta_3 \times \mbox{victim\_sex}_i + \beta_4 \times \mbox{city}_i + \beta_5 \times \mbox{year}_i\\
\beta_0 &\sim \mbox{Normal}(0, 2.5) \\
\beta_1 &\sim \mbox{Normal}(0, 2.5) \\
\beta_2 &\sim \mbox{Normal}(0, 2.5) \\
\beta_3 &\sim \mbox{Normal}(0, 2.5) \\
\beta_4 &\sim \mbox{Normal}(0, 2.5) \\
\beta_5 &\sim \mbox{Normal}(0, 2.5)
\end{align*}

@@

### Idealized Survey Objectives
The objective of our survey is to collect information from Chicago Police Department and Los Angeles Police Department about their education background, reason for closing a homicide case without arrest, difficult homicide case they have done, how many cases they have to cover at a time, etc.

### Sampling Approach
- what is the population, frame, and sample;
- how is the sample recruited;
- what sampling approach is taken, and what are some of the trade-offs of this; 
- how is non-response handled;

The sampling approach we plan to take is stratified sampling, which is a type of probabilistic sampling [@citetellingstories]

### Respondent Recruitment

@@

>>>>

The raw dataset we retrieved from The Washington Post using a script that downloads the CSV file from their GitHub repository contains 52179 observations with one observation being a homicide case. However, since we narrowed our scope to focus on the two of the most populated US cities, Los Angeles and Chicago, the number of observations in our cleaned dataset ended up being 6307. Our data look as follows:

@@

@tbl-dataset-preview and @tbl-dataset-statistics indicates our variables of interest, which are the following: victim_race, victim_age, victim_sex, city, disposition, year, month, and arrest_was_not_made. victim_race represents the race of the homicide victim, which can be "White" (376 observations), "Hispanic" (1763 observations), "Black" (4063 observations), "Asian" (43 observations), and "Other" (62 observations). victim_age signifies the age of the homicide victim at the time of their death and is defined as a whole number. @tbl-dataset-statistics reveals that the mean victim_sex is the sex of the homicide victim where they are either identified as a "female" or "male". The city variable defines the city the victim is reported to have been found in. The disposition variable is the specific status of a homicide case where a case can fall in either of the following three status: "Closed by arrest", "Open/No arrest", and "Closed without arrest". The year variable represents a year from 2010 to 2017 that indicates the year the homicide took place. Following this, the month variable represents the month a homicide took place. arrest_was_not_made is a variable that was constructed based on the disposition variable with "Closed by arrest" being converted to a 0 and "Open/No arrest" and "Closed without arrest" being converted to a 1. arrest_was_not_made indicates the status of a homicide case as either being unsolved, which is denoted by a 1, and solved, which is denoted by a 0.

@@

The Federal Bureau of Investigation (FBI) has a program called the Uniform Crime Reporting (UCR) Program to generate statistics for the public [@citeFBIinfo]. The FBI originally had a system under the UCR called Summary Reporting System (SRS), which obtained details about different crimes taking place from law enforcement agencies nationwide such as victim information [@citeFBIinfo]. However, the SRS was replaced with a new system called the National Incident-Based Reporting System (NIBR) in 2021 that obtained more details about various crimes [@citeFBIinfo]. For The Los Angeles Police Department, after a homicide case occurs, the investigating team handwrites information into physical crime reports with details like the type of crime, the premise the crime occured at, and the age and ethnicity of the victim [@citeLApolice]. The crime reports are then transcribed into a digital format and then sent to the SRS monthly [@citeLApolice]. After a homicide occurs in Chicago, The Chicago Police Department uses a system called the Chicago Police Department's CLEAR (Citizen Law Enforcement Analysis and Reporting) system to report on details such as the victim, if an arrest was made or not, and the location the crime took [@citeChicagoPolice]. This information is then reported to the FBI under the UCR program.

Reporters from the Washington Post then obtained the data from the FBI specifically on homicides from 50 large US cities from 2007 to 2010, which can be access through the UCR publications page on the FBI website [@datahomicides]. They also obtain data about homicide counts and closure rates through papers and compare these values with the ones from the FBI dataset for accuracy [@datahomicides]. The Washington Post would also use public records like medical examiner reports, death certificates, and court records, to fill in any missing information since some departments only report partial information to the FBI [@datahomicides]. The Washington Post defined cases to be closed without arrest when they are reported by the police as "exceptionally cleared" where there is evidence of who the perpetrator is but an arrest is not possible because of reason like they has died [@datahomicides]. They also define cases to be closed by arrest if the police reported it to be and other cases are defined to be open/no arrest [@datahomicides].

## Outcome variable
The outcome variable we are interested in looking at with our model and our analysis is the arrest_was_not_made variable. We use this variable to compare homicide case characteristics of solved and unsolved homicides.

## Predictor variables
The predictor variables for our model are the following: victim_race, victim_age, victim_sex, city, and year. The variables, disposition and month are not predictor variables in our model but they are used to investigate trends between homicide case information and the status of the homicide being solved or unsolved.

@@

## Model set-up
With our model, we will make the assumption that there is a relationship between homicide case information like the victim's race, victim's age, victim's sex, the city, and the year with a homicide case being unsolved. We also assume that the predictor variables are independent from one another, which we check in @sec-model-details using variance inflation factor (VIF) and it indicates the predictors are not highly correlated with each other. We define our model as follows:
\begin{align*}
y_i|\pi_i &\sim \mbox{Bern}(\pi_i) \\
\mbox{logit}(\pi_i) &= \beta_0 + \beta_1 \times \mbox{victim\_race}_i + \beta_2 \times \mbox{victim\_age}_i + \beta_3 \times \mbox{victim\_sex}_i + \beta_4 \times \mbox{city}_i + \beta_5 \times \mbox{year}_i\\
\beta_0 &\sim \mbox{Normal}(0, 2.5) \\
\beta_1 &\sim \mbox{Normal}(0, 2.5) \\
\beta_2 &\sim \mbox{Normal}(0, 2.5) \\
\beta_3 &\sim \mbox{Normal}(0, 2.5) \\
\beta_4 &\sim \mbox{Normal}(0, 2.5) \\
\beta_5 &\sim \mbox{Normal}(0, 2.5)
\end{align*}

@@

### Idealized Survey Objectives
The objective of our survey is to gain insight about the investigators from US police departments and their experience with dealing with homicide cases to take into account into any potential factors about police officers that could potentially impact if a homicide case ends up unsolved. In the following idealized methodology, we will cover our sampling approach, respondent recruitment, data validation, the design of our survey, limitations of our survey, and the survey questions themselves.

### Sampling Approach
- what is the population, frame, and sample;
- what sampling approach is taken, and what are some of the trade-offs of this; 
- how is non-response handled;

Our target population are all investigators from US police departments [@citetellingstories].

Chicago and Los Angeles Police Departments 
The sampling approach we plan to take is stratified sampling, which is a type of probabilistic sampling [@citetellingstories]

### Respondent Recruitment
- how is the sample recruited;

@@

about their education background, reason for closing a homicide case without arrest, difficult homicide case they have done, how many cases they have to cover at a time, etc.
